---
title: "Comparing with mean (with signal) Recover"
author: "Yuxin Zou"
date: 2018-05-16
output: 
  html_document:
    code_folding: hide
---
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the R version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r, echo=FALSE}
# TEMPORARY.
knitr::opts_chunk$set(eval = TRUE)
```

```{r, echo=FALSE}
library(mashr)
library(corrplot); library(ggplot2)
source('../code/sim_mean_sig.R')
source('../code/MashSource.R')
```

The data contains 10 conditions with 10% non-null samples. For the non-null samples, it has equal effects in the first c conditions. 

Let L be the contrast matrix that subtract mean from each sample.

$$\hat{\delta}_{j}|\delta_{j} \sim N(\delta_{j}, \frac{1}{2}LL')$$
90% of the true deviations are 0. 10% of the deviation $\delta_{j}$ has correlation that the first c conditions are negatively correlated with the rest conditions.


We set $c = 2$.
```{r data}
set.seed(1)
R = 10
C = 2
data = sim.mean.sig(nsamp=10000, ncond=C)
```

# Discard last column

## Mash contrast model
```{r}
L = matrix(-1/R, R, R)
L[cbind(1:R,1:R)] = (R-1)/R
row.names(L) = colnames(data$Chat)
L.10 = L[1:(R-1),]
mash_data = mash_set_data(Bhat=data$Chat, Shat=data$Shat)
mash_data_L.10 = mash_set_data_contrast(mash_data, L.10)
```

```{r}
U.c = cov_canonical(mash_data_L.10)

# data driven
# select max
m.1by1 = mash_1by1(mash_data_L.10)
strong = get_significant_results(m.1by1,0.05)
# center Z
mash_data_L.center = mash_data_L.10
mash_data_L.center$Bhat = mash_data_L.10$Bhat/mash_data_L.10$Shat # obtain z
mash_data_L.center$Shat = matrix(1, nrow(mash_data_L.10$Bhat),ncol(mash_data_L.10$Bhat))
mash_data_L.center$Bhat = apply(mash_data_L.center$Bhat, 2, function(x) x - mean(x))
U.pca = cov_pca(mash_data_L.center,2,strong)
U.ed = cov_ed(mash_data_L.center, U.pca, strong)

mashcontrast.model.10 = mash(mash_data_L.10, c(U.c, U.ed), algorithm.version = 'R', verbose = FALSE)
```
Using `mashcommonbaseline`, there are `r length(get_significant_results(mashcontrast.model.10))` discoveries. The covariance structure found here is:
```{r}
barplot(get_estimated_pi(mashcontrast.model.10),las = 2, cex.names = 0.7)
```
The correlation for PCA1 is:
```{r, echo=FALSE, fig.width=3, fig.height=3,fig.align='center'}
x           <- cov2cor(mashcontrast.model.10$fitted_g$Ulist[["ED_PCA_1"]])
x[x > 1]    <- 1
x[x < -1]   <- -1
colnames(x) <- colnames(get_lfsr(mashcontrast.model.10))
rownames(x) <- colnames(x)
corrplot.mixed(x,upper='color',cl.lim=c(-1,1), upper.col=colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(40),
               title='PCA1',mar=c(0,0,1.5,0))
```

Recover the last column
```{r}
mashcontrast.model.10.full = mashcontrast.model.10
mashcontrast.model.10.full$result = mash_compute_posterior_matrices(g = mashcontrast.model.10, data = mash_data_L.10, algorithm.version = 'R', recover=TRUE)
```

There are `r length(get_significant_results(mashcontrast.model.10.full))` discoveries.

* Original estimates
```{r echo=FALSE}
# raw effect
mash_data_L = mash_set_data_contrast(mash_data, L)
stronggene = data.frame(mash_data_L$Bhat[ get_significant_results(mashcontrast.model.10.full)[1],])
colnames(stronggene) = 'EffectSize'
stronggene$Group = row.names(stronggene)
stronggene$Group <- factor(stronggene$Group, levels = paste0('condition_',1:10))
stronggene$se = mash_data_L$Shat[ get_significant_results(mashcontrast.model.10.full)[1],]
ggplot(stronggene, aes(y = EffectSize, x = Group, color=Group)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=EffectSize-1.96*se, ymax=EffectSize+1.96*se), width=0.4) + 
  theme_bw(base_size=12) + coord_flip()
```

* `MASH` estimates

```{r echo=FALSE}
# Improved
stronggenePost = data.frame(mashcontrast.model.10.full$result$PosteriorMean[get_significant_results(mashcontrast.model.10.full)[1],])
colnames(stronggenePost) = 'EffectSize'
stronggenePost$Group = row.names(stronggenePost)
stronggenePost$se = mashcontrast.model.10.full$result$PosteriorSD[get_significant_results(mashcontrast.model.10.full)[1],]
ggplot(stronggenePost, aes(y = EffectSize, x = Group, color=Group)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=EffectSize-1.96*se, ymax=EffectSize+1.96*se), width=0.4) + 
  theme_bw(base_size=12) + coord_flip()
```

## Subtract mean directly

If we subtract the mean from the data directly
$$Var(\hat{c}_{j,r}-\bar{\hat{c}_{j}}) = \frac{1}{2} - \frac{1}{2R}$$
```{r}
Indep.data.10 = mash_set_data(Bhat = mash_data_L.10$Bhat,
                           Shat = matrix(sqrt(0.5-1/(2*R)), nrow(data$Chat), R-1))

Indep.model.10 = mash(Indep.data.10, c(U.c, U.ed), algorithm.version = 'R', verbose = FALSE)
```
There are `r length(get_significant_results(Indep.model.10))` discoveries. The covariance structure found here is:
```{r}
barplot(get_estimated_pi(Indep.model.10),las = 2, cex.names = 0.7)
```

The correlation for PCA2 and tPCA is:
```{r, echo=FALSE, fig.width=8, fig.height=3,fig.align='center'}
par(mfrow=c(1,2))
x           <- cov2cor(Indep.model.10$fitted_g$Ulist[["ED_PCA_2"]])
x[x > 1]    <- 1
x[x < -1]   <- -1
colnames(x) <- seq(1,9)
rownames(x) <- seq(1,9)
corrplot.mixed(x,upper='color',cl.lim=c(-1,1), upper.col=colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(40),
               title='PCA2',mar=c(0,0,1.5,0))

x           <- cov2cor(Indep.model.10$fitted_g$Ulist[["ED_tPCA"]])
x[x > 1]    <- 1
x[x < -1]   <- -1
colnames(x) <- seq(1,9)
rownames(x) <- seq(1,9)
corrplot.mixed(x,upper='color',cl.lim=c(-1,1), upper.col=colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(40),
               title='tPCA',mar=c(0,0,1.5,0))
par(mfrow=c(1,1))
```

Recover the last column
```{r}
Indep.model.10.full = Indep.model.10
Indep.model.10.full$result = mash_compute_posterior_matrices(g = Indep.model.10, data = Indep.data.10, algorithm.version = 'R', recover=TRUE)
```

There are `r length(get_significant_results(Indep.model.10.full))` discoveries.

## Miscalculation of variance of mean

We try an example with the miscalculated variance of mean. The reason to include this example is that the variance of median is hard to compute (not iid data), if we subtract median from the samples directly. We want to test whether the misspecified variance could influence the result.

The following model is fitted under miscalculation of $Var(\hat{c}_{j,r}-\bar{\hat{c}_{j}}) = \frac{1}{2}$.

```{r, echo=FALSE}
Mis.data.10 = mash_set_data(Bhat = mash_data_L.10$Bhat, Shat=sqrt(0.5))
# data driven
# select max
m.1by1 = mash_1by1(Mis.data.10)
strong = get_significant_results(m.1by1,0.05)
# center Z
mash_data_L.center = Mis.data.10
mash_data_L.center$Bhat = Mis.data.10$Bhat/Mis.data.10$Shat # obtain z
mash_data_L.center$Shat = matrix(1, nrow(Mis.data.10$Bhat),ncol(Mis.data.10$Bhat))
mash_data_L.center$Bhat = apply(mash_data_L.center$Bhat, 2, function(x) x - mean(x))
U.pca = cov_pca(mash_data_L.center,2,strong)
U.ed = cov_ed(mash_data_L.center, U.pca, strong)

Mis.model.10 = mash(Mis.data.10, c(U.c, U.ed), algorithm.version = 'R', verbose = FALSE)
```
There are `r length(get_significant_results(Mis.model.10))` discoveries. The covariance structure found here is:
```{r, echo=FALSE}
barplot(get_estimated_pi(Mis.model.10),las = 2, cex.names = 0.7)
```

Recover the last column
```{r}
Mis.model.10.full = Mis.model.10
Mis.model.10.full$result = mash_compute_posterior_matrices(g = Mis.model.10, data = Mis.data.10, algorithm.version = 'R', recover=TRUE)
```

There are `r length(get_significant_results(Mis.model.10.full))` discoveries.


# Discard the first column

The data was generated with signals in the first c conditions ($c_{j,1}, \cdots, c_{j,c}$). The contrast matrix L used here discards the last condition among R conditions. The deviations are $\hat{c}_{j,1} - \bar{\hat{c}_{j}}, \hat{c}_{j,2} - \bar{\hat{c}_{j}}, \cdots, \hat{c}_{j,R-1} - \bar{\hat{c}_{j}}$.

However, the contrast matrix L can discard any deviation from $\hat{c}_{j,1} - \bar{\hat{c}_{j}}, \cdots, \hat{c}_{j,R} - \bar{\hat{c}_{j}}$. The choice of the discarded deviation could influence the result.

We run the same model with L that discard the first deviation.

## Mash contrast model

```{r}
L.1 = L[2:R,]
mash_data_L.1 = mash_set_data_contrast(mash_data, L.1)
```

```{r}
U.c = cov_canonical(mash_data_L.1)

# data driven
# select max
m.1by1 = mash_1by1(mash_data_L.1)
strong = get_significant_results(m.1by1,0.05)
# center Z
mash_data_L.center = mash_data_L.1
mash_data_L.center$Bhat = mash_data_L.1$Bhat/mash_data_L.1$Shat # obtain z
mash_data_L.center$Shat = matrix(1, nrow(mash_data_L.1$Bhat),ncol(mash_data_L.1$Bhat))
mash_data_L.center$Bhat = apply(mash_data_L.center$Bhat, 2, function(x) x - mean(x))
U.pca = cov_pca(mash_data_L.center,2,strong)
U.ed = cov_ed(mash_data_L.center, U.pca, strong)

mashcontrast.model.1 = mash(mash_data_L.1, c(U.c, U.ed), algorithm.version = 'R', verbose = FALSE)
```
Using `mashcommonbaseline` model, there are `r length(get_significant_results(mashcontrast.model.1))` discoveries. The covariance structure found here is:
```{r}
barplot(get_estimated_pi(mashcontrast.model.1),las = 2, cex.names = 0.7)
```
The correlation PCA 1 is:
```{r, echo=FALSE, fig.width=3, fig.height=3,fig.align='center'}
x           <- cov2cor(mashcontrast.model.1$fitted_g$Ulist[["ED_PCA_1"]])
x[x > 1]    <- 1
x[x < -1]   <- -1
colnames(x) <- colnames(get_lfsr(mashcontrast.model.1))
rownames(x) <- colnames(x)
corrplot.mixed(x,upper='color',cl.lim=c(-1,1), upper.col=colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(40),
               title='PCA1',mar=c(0,0,1.5,0))
```

Recover the first column
```{r}
mashcontrast.model.1.full = mashcontrast.model.1
mashcontrast.model.1.full$result = mash_compute_posterior_matrices(g = mashcontrast.model.1, data = mash_data_L.1, algorithm.version = 'R', recover=TRUE)
```

There are `r length(get_significant_results(mashcontrast.model.1.full))` discoveries.

## Subtract mean directly

```{r}
Indep.data.1 = mash_set_data(Bhat = mash_data_L.1$Bhat,
                           Shat = matrix(sqrt(0.5-1/(R*2)), nrow(data$Chat), R-1))

Indep.model.1 = mash(Indep.data.1, c(U.c, U.ed), algorithm.version = 'R', verbose = FALSE)
```
For `mashIndep` model, there are `r length(get_significant_results(Indep.model.1))` discoveries. The covariance structure found here is:
```{r}
barplot(get_estimated_pi(Indep.model.1),las = 2, cex.names = 0.7)
```

The correlation for PCA2 and tPCA is:
```{r, echo=FALSE, fig.width=8, fig.height=3,fig.align='center'}
par(mfrow=c(1,2))
x           <- cov2cor(Indep.model.1$fitted_g$Ulist[["ED_PCA_2"]])
x[x > 1]    <- 1
x[x < -1]   <- -1
colnames(x) <- seq(1,9)
rownames(x) <- seq(1,9)
corrplot.mixed(x,upper='color',cl.lim=c(-1,1), upper.col=colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(40),
               title='PCA2',mar=c(0,0,1.5,0))

x           <- cov2cor(Indep.model.1$fitted_g$Ulist[["ED_tPCA"]])
x[x > 1]    <- 1
x[x < -1]   <- -1
colnames(x) <- seq(1,9)
rownames(x) <- seq(1,9)
corrplot.mixed(x,upper='color',cl.lim=c(-1,1), upper.col=colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(40),
               title='tPCA',mar=c(0,0,1.5,0))
par(mfrow=c(1,1))
```

Recover the first column
```{r}
Indep.model.1.full = Indep.model.1
Indep.model.1.full$result = mash_compute_posterior_matrices(g = Indep.model.1, data = Indep.data.1, algorithm.version = 'R', recover=TRUE)
```

There are `r length(get_significant_results(Indep.model.1.full))` discoveries.

## Miscalculation of variance of mean

The following model is fitted under miscalculation of $Var(\hat{c}_{j,r}-\bar{\hat{c}_{j}}) = \frac{1}{2}$.

```{r, echo=FALSE}
Mis.data.1 = mash_set_data(Bhat = mash_data_L.1$Bhat, Shat=sqrt(0.5))
# data driven
# select max
m.1by1 = mash_1by1(Mis.data.1)
strong = get_significant_results(m.1by1,0.05)
# center Z
mash_data_L.center = Mis.data.1
mash_data_L.center$Bhat = Mis.data.1$Bhat/Mis.data.1$Shat # obtain z
mash_data_L.center$Shat = matrix(1, nrow(Mis.data.1$Bhat),ncol(Mis.data.1$Bhat))
mash_data_L.center$Bhat = apply(mash_data_L.center$Bhat, 2, function(x) x - mean(x))
U.pca = cov_pca(mash_data_L.center,2,strong)
U.ed = cov_ed(mash_data_L.center, U.pca, strong)

Mis.model.1 = mash(Mis.data.1, c(U.c, U.ed), algorithm.version = 'R', verbose = FALSE)
```
There are `r length(get_significant_results(Mis.model.1))` discoveries. The covariance structure found here is:
```{r, echo=FALSE}
barplot(get_estimated_pi(Mis.model.1),las = 2, cex.names = 0.7)
```

Recover the first column
```{r}
Mis.model.1.full = Mis.model.1
Mis.model.1.full$result = mash_compute_posterior_matrices(g = Mis.model.1, data = Mis.data.1, algorithm.version = 'R', recover=TRUE)
```

There are `r length(get_significant_results(Mis.model.10.full))` discoveries.


# Compare models

The RRMSE plot:
```{r}
delta.10 = data$C - rowMeans(data$C)
deltahat.10 = data$Chat - rowMeans(data$Chat)

delta.1 = delta.10[, c(2:10, 1)]
deltahat.1 = deltahat.10[, c(2:10, 1)]

barplot(c(sqrt(mean((delta.10 - mashcontrast.model.10.full$result$PosteriorMean)^2)/mean((delta.10 - deltahat.10)^2)), 
          sqrt(mean((delta.1 - mashcontrast.model.1.full$result$PosteriorMean)^2)/mean((delta.1 - deltahat.1)^2)), 
          sqrt(mean((delta.10 - Indep.model.10.full$result$PosteriorMean)^2)/mean((delta.10 - deltahat.10)^2)),
          sqrt(mean((delta.1 - Indep.model.1.full$result$PosteriorMean)^2)/mean((delta.1 - deltahat.1)^2)), 
          sqrt(mean((delta.10 - Mis.model.10.full$result$PosteriorMean)^2)/mean((delta.10 - deltahat.10)^2)), 
          sqrt(mean((delta.1 - Mis.model.1.full$result$PosteriorMean)^2)/mean((delta.1 - deltahat.1)^2))), ylim=c(0,0.2), names.arg = c('mashcommon.10','mashcommon.1','mash.indep.10', 'mash.indep.1', 'mis.10', 'mis.1'), ylab='RRMSE')
```

We check the False Positive Rate and True Positive Rate. 
$$FPR = \frac{|N\cap S|}{|N|} \quad TPR = \frac{|CS\cap S|}{|T|} $$

```{r}
sign.test.mash.10 = as.matrix(delta.10)*mashcontrast.model.10.full$result$PosteriorMean
sign.test.Indep.10 = as.matrix(delta.10)*Indep.model.10.full$result$PosteriorMean
sign.test.Mis.10 = as.matrix(delta.10)*Mis.model.10.full$result$PosteriorMean
sign.test.mash.1 = as.matrix(delta.1)*mashcontrast.model.1.full$result$PosteriorMean
sign.test.Indep.1 = as.matrix(delta.1)*Indep.model.1.full$result$PosteriorMean
sign.test.Mis.1 = as.matrix(delta.1)*Mis.model.1.full$result$PosteriorMean


thresh.seq = seq(0, 1, by=0.0005)[-1]
mashcontrast.1 = matrix(0,length(thresh.seq), 2)
Indep.1 = matrix(0,length(thresh.seq), 2)
Mis.1 = matrix(0,length(thresh.seq), 2)
mashcontrast.10 = matrix(0,length(thresh.seq), 2)
Indep.10 = matrix(0,length(thresh.seq), 2)
Mis.10 = matrix(0,length(thresh.seq), 2)
colnames(mashcontrast.1) = colnames(Mis.10) = colnames(Mis.1) = c('TPR', 'FPR')
colnames(Indep.1) = c('TPR', 'FPR')
colnames(mashcontrast.10) = c('TPR', 'FPR')
colnames(Indep.10) = c('TPR', 'FPR')
for(t in 1:length(thresh.seq)){
  mashcontrast.1[t,] = c(sum(sign.test.mash.1>0 & mashcontrast.model.1.full$result$lfsr <= thresh.seq[t])/sum(delta.1!=0), sum(delta.1==0 & mashcontrast.model.1.full$result$lfsr <=thresh.seq[t])/sum(delta.1==0))
  
  Indep.1[t,] = c(sum(sign.test.Indep.1>0& Indep.model.1.full$result$lfsr <=thresh.seq[t])/sum(delta.1!=0),  sum(delta.1==0& Indep.model.1.full$result$lfsr <=thresh.seq[t])/sum(delta.1==0))
  
  Mis.1[t,] = c(sum(sign.test.Mis.1>0 & Mis.model.1.full$result$lfsr <= thresh.seq[t])/sum(delta.1!=0), sum(delta.1==0 & Mis.model.1.full$result$lfsr <=thresh.seq[t])/sum(delta.1==0))
  
  mashcontrast.10[t,] = c(sum(sign.test.mash.10>0 & mashcontrast.model.10.full$result$lfsr <= thresh.seq[t])/sum(delta.10!=0), sum(delta.10==0 & mashcontrast.model.10.full$result$lfsr <=thresh.seq[t])/sum(delta.10==0))
  
  Indep.10[t,] = c(sum(sign.test.Indep.10>0& Indep.model.10.full$result$lfsr <=thresh.seq[t])/sum(delta.10!=0),  sum(delta.10==0& Indep.model.10.full$result$lfsr <=thresh.seq[t])/sum(delta.10==0))
  
  Mis.10[t,] = c(sum(sign.test.Mis.10>0& Mis.model.10.full$result$lfsr <=thresh.seq[t])/sum(delta.10!=0),  sum(delta.10==0& Mis.model.10.full$result$lfsr <=thresh.seq[t])/sum(delta.10==0))
}
```

```{r, echo=FALSE}
{plot(mashcontrast.1[,'FPR'], mashcontrast.1[,'TPR'], col='red',type='l',xlab = 'FPR', ylab='TPR', main='True Positive vs False Positive')
lines(mashcontrast.10[,'FPR'], mashcontrast.10[,'TPR'], col='green')
lines(Indep.1[,'FPR'], Indep.1[,'TPR'], col='blue')
lines(Indep.10[,'FPR'], Indep.10[,'TPR'])
lines(Mis.1[,'FPR'], Mis.1[,'TPR'], col='brown')
lines(Mis.10[,'FPR'], Mis.10[,'TPR'], col='cyan')
legend('bottomright', c('mashcommon.1','mashcommon.10', 'mash.indep.1', 'mash.indep.10', 'mis.1', 'mis.10'),col=c('red','green', 'blue', 'black', 'brown', 'cyan'),lty=c(1,1,1,1,1,1))}
```

The `mashcommonbaseline` model performs better than `mash.indep` model. The choice of the discarded column has larger effect for `mash.indep` model.

# Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
