---
title: "MASH-Flash 1 factor"
author: "Yuxin Zou"
date: 2018-01-22
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the R version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r R-version, echo=FALSE, results='asis'}
```

```{r, echo=FALSE}
# TEMPORARY.
knitr::opts_chunk$set(eval = TRUE)
```


```{r echo=FALSE}
library(flashr)
library(mashr)
library(corrplot)
source('../code/sim.R')
```

# Generate data

```{r}
set.seed(50)
data = sim(nsamp=1000, err_sd = 0.05)
```

* `flashr` Factorization

```{r eval=FALSE}
z = data$Bhat/data$Shat
# center
Z.center = apply(z, 2, function(x) x - mean(x))

fdata = flash_set_data(Z.center)
fm = flash(fdata,greedy = TRUE, backfit = TRUE)

```

The factor structures:

```{r echo=FALSE}
factors = flash_get_ldf(fm)$f
pve.order = order(flash_get_pve(fm), decreasing = TRUE)

par(mfrow=c(2,2))
for(i in pve.order){
  barplot(factors[,i], main=paste0('Factor ',i, ' pve= ', round(flash_get_pve(fm)[i],3)), las=2, cex.names = 0.7)
}
par(mfrow=c(1,1))
```

# Apply `mash` without factor rank 1 covariance

```{r echo=FALSE}
mash_data = mash_set_data(Bhat = data$Bhat, Shat = data$Shat)
U.c = cov_canonical(mash_data)
U.flash = c(list("tFlash" = t(flash_get_lf(fm)) %*% flash_get_lf(fm) / nrow(mash_data$Bhat)))

U.ed = cov_ed(mash_data, U.flash)
mash_model = mash(mash_data, c(U.c,U.ed), algorithm.version = 'R', verbose = FALSE)
barplot(get_estimated_pi(mash_model), las=2, cex.names = 0.7)
```

The `mash` model puts the majority weights on `tFlash`. There are `r length(get_significant_results(mash_model))` siginificant findings. The log likelihood is `r get_loglik(mash_model)`.

# Apply `mash` with factor rank 1 covariance

```{r echo=FALSE}
U.flash.1 = c(mashr::cov_from_factors(t(factors), "Flash"),
            list("tFlash" = t(flash_get_lf(fm)) %*% flash_get_lf(fm) / nrow(mash_data$Bhat)))

U.ed.1 = cov_ed(mash_data, U.flash.1)
mash_model.1 = mash(mash_data, c(U.c,U.ed.1), algorithm.version = 'R', verbose = FALSE)
barplot(get_estimated_pi(mash_model.1), las=2, cex.names = 0.7)
```
The `mash` model identifies the covariacne structure correctly. There are `r length(get_significant_results(mash_model.1))` siginificant findings. The log likelihood is `r get_loglik(mash_model.1)`.

With the correct covariance struture, the `mash` model has a much higher likelihood value. Adding rank 1 factor covariance improve the fitting.

# Session information

<!-- Insert the session information into the document -->
```{r session-info}
```

