---
title: "MeanSignal_CondnMean"
author: "Yuxin Zou"
date: 2018-5-16
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

```{r code-version, echo=FALSE, results='asis'}
```

```{r, echo=FALSE}
library(mashr)
library(corrplot); library(ggplot2)
source('../code/sim_mean_sig.R')
source('../code/MashSource.R')
```

The data contains 10 conditions with 10% non-null samples. For the non-null samples, it has equal effects in the first c conditions. 

Let L be the contrast matrix that substract mean from each sample.

$$\hat{\delta}_{j}|\delta_{j} \sim N(\delta_{j}, \frac{1}{2}LL')$$
90% of the true deviations are 0. 10% of the deviation $\delta_{j}$ has correlation that the first c conditions are negatively correlated with the rest conditions.

We set $c = 2$.
```{r, echo=FALSE}
set.seed(1)
R = 10
C = 2
data = sim.mean.sig(nsamp=10000, ncond=C)
```

# Discard last col
```{r, echo=FALSE}
L = matrix(-1/R, R, R)
diag(L) = (R-1)/R
row.names(L) = colnames(data$Chat)
mash_data = mash_set_data(Bhat=data$Chat, Shat=data$Shat)
```

## Mash

```{r, echo=FALSE}
L.10 = L[1:R-1,]
mash_data_L = mash_set_data_contrast(mash_data, L.10)
U.c = cov_canonical(mash_data_L)

# data driven
# select max
m.1by1 = mash_1by1(mash_data_L, alpha=0)
strong = get_significant_results(m.1by1,0.05)
# center Z
mash_data.center = mash_data_L
mash_data.center$Bhat = mash_data_L$Bhat/mash_data_L$Shat # obtain z
mash_data.center$Shat = matrix(1, nrow(mash_data_L$Bhat),ncol(mash_data_L$Bhat))
mash_data.center$Bhat = apply(mash_data.center$Bhat, 2, function(x) x - mean(x))
U.pca = cov_pca(mash_data.center,2,strong)
U.ed = cov_ed(mash_data.center, U.pca, strong)

mashcontrast.model.10 = mash(mash_data_L, c(U.c, U.ed), algorithm.version = 'R')
```

Using `mashcommonbaseline`, there are `r length(get_significant_results(mashcontrast.model.10))` discoveries. The covariance structure found here is:
```{r, echo=FALSE}
barplot(get_estimated_pi(mashcontrast.model.10),las = 2, cex.names = 0.7)
```

Recover the last column
```{r}
mashcontrast.model.10.full = mashcontrast.model.10
mashcontrast.model.10.full$result = mash_compute_posterior_matrices(g = mashcontrast.model.10, data = mash_data_L, algorithm.version = 'R', recover=TRUE)
```

There are `r length(get_significant_results(mashcontrast.model.10.full))` discoveries.

## Bovy
```{r, echo=FALSE}
mash_data_L = mash_set_data_contrast(mash_data, L)
U.c = cov_canonical(mash_data_L)

# data driven
# select max
m.1by1 = mash_1by1(mash_data_L, alpha=0)
strong = get_significant_results(m.1by1,0.05)
# center Z
mash_data.center = mash_data_L
mash_data.center$Bhat = mash_data_L$Bhat/mash_data_L$Shat # obtain z
mash_data.center$Shat = matrix(1, nrow(mash_data_L$Bhat),ncol(mash_data_L$Bhat))
mash_data.center$Bhat = apply(mash_data.center$Bhat, 2, function(x) x - mean(x))
U.pca = cov_pca(mash_data.center,2,strong)
U.ed = cov_ed(mash_data.center, U.pca, strong)

L.10 = L[1:(R-1),]
mash_data_L_10 = mash_set_data_contrast(mash_data, L.10)
mash_data_L_10$Shat_alpha = mash_data_L$Shat_alpha
Bovy.model.10 = mash(mash_data_L_10, c(U.c, U.ed), algorithm.version = 'R', Bovy=TRUE)
```

Using `Bovy`, there are `r length(get_significant_results(Bovy.model.10))` discoveries. The covariance structure found here is:
```{r, echo=FALSE}
barplot(get_estimated_pi(Bovy.model.10),las = 2, cex.names = 0.7)
```

The row means of the posterior mean matrix are not 0.
```{r}
rowMeans(Bovy.model.10$result$PosteriorMean)[1:6]
```

* Original estimates
```{r echo=FALSE}
# raw effect
mash_data_L = mash_set_data_contrast(mash_data, L)
stronggene = data.frame(mash_data_L$Bhat[9211,])
colnames(stronggene) = 'EffectSize'
stronggene$Group = row.names(stronggene)
stronggene$Group <- factor(stronggene$Group, levels = paste0('condition_',1:10))
stronggene$se = mash_data_L$Shat[9211,]
ggplot(stronggene, aes(y = EffectSize, x = Group, color=Group)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=EffectSize-1.96*se, ymax=EffectSize+1.96*se), width=0.4) + 
  theme_bw(base_size=12) + coord_flip()
```

* `MASH` estimates

```{r echo=FALSE}
# Improved
stronggenePost = data.frame(mashcontrast.model.10.full$result$PosteriorMean[9211,])
colnames(stronggenePost) = 'EffectSize'
stronggenePost$Group = row.names(stronggenePost)
stronggenePost$se = mashcontrast.model.10.full$result$PosteriorSD[9211,]
ggplot(stronggenePost, aes(y = EffectSize, x = Group, color=Group)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=EffectSize-1.96*se, ymax=EffectSize+1.96*se), width=0.4) + 
  theme_bw(base_size=12) + coord_flip()
```

* `Bovy` estimates

```{r echo=FALSE}
# Improved
stronggenePost = data.frame(Bovy.model.10$result$PosteriorMean[9211,])
colnames(stronggenePost) = 'EffectSize'
stronggenePost$Group = row.names(stronggenePost)
stronggenePost$se = Bovy.model.10$result$PosteriorSD[9211,]
ggplot(stronggenePost, aes(y = EffectSize, x = Group, color=Group)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=EffectSize-1.96*se, ymax=EffectSize+1.96*se), width=0.4) + 
  theme_bw(base_size=12) + coord_flip()
```

```{r}
mean((Bovy.model.10$result$PosteriorMean - mashcontrast.model.10.full$result$PosteriorMean)^2)
```

# Discard first col

## Mash

```{r, echo=FALSE}
L.1 = L[2:R,]
mash_data_L = mash_set_data_contrast(mash_data, L.1)
U.c = cov_canonical(mash_data_L)

# data driven
# select max
m.1by1 = mash_1by1(mash_data_L, alpha=0)
strong = get_significant_results(m.1by1,0.05)
# center Z
mash_data.center = mash_data_L
mash_data.center$Bhat = mash_data_L$Bhat/mash_data_L$Shat # obtain z
mash_data.center$Shat = matrix(1, nrow(mash_data_L$Bhat),ncol(mash_data_L$Bhat))
mash_data.center$Bhat = apply(mash_data.center$Bhat, 2, function(x) x - mean(x))
U.pca = cov_pca(mash_data.center,2,strong)
U.ed = cov_ed(mash_data.center, U.pca, strong)

mashcontrast.model.1 = mash(mash_data_L, c(U.c, U.ed), algorithm.version = 'R')
```

Using `mashcommonbaseline`, there are `r length(get_significant_results(mashcontrast.model.1))` discoveries. The covariance structure found here is:
```{r, echo=FALSE}
barplot(get_estimated_pi(mashcontrast.model.1),las = 2, cex.names = 0.7)
```

Recover the last column
```{r}
mashcontrast.model.1.full = mashcontrast.model.1
mashcontrast.model.1.full$result = mash_compute_posterior_matrices(g = mashcontrast.model.1, data = mash_data_L, algorithm.version = 'R', recover=TRUE)
```

There are `r length(get_significant_results(mashcontrast.model.1.full))` discoveries.

## Bovy
```{r, echo=FALSE}
mash_data_L = mash_set_data_contrast(mash_data, L)
U.c = cov_canonical(mash_data_L)

# data driven
# select max
m.1by1 = mash_1by1(mash_data_L, alpha=0)
strong = get_significant_results(m.1by1,0.05)
# center Z
mash_data.center = mash_data_L
mash_data.center$Bhat = mash_data_L$Bhat/mash_data_L$Shat # obtain z
mash_data.center$Shat = matrix(1, nrow(mash_data_L$Bhat),ncol(mash_data_L$Bhat))
mash_data.center$Bhat = apply(mash_data.center$Bhat, 2, function(x) x - mean(x))
U.pca = cov_pca(mash_data.center,2,strong)
U.ed = cov_ed(mash_data.center, U.pca, strong)

L.1 = L[2:R,]
mash_data_L_1 = mash_set_data_contrast(mash_data, L.1)
mash_data_L_1$Shat_alpha = mash_data_L$Shat_alpha
Bovy.model.1 = mash(mash_data_L_1, c(U.c, U.ed), algorithm.version = 'R', Bovy=TRUE)
```

Using `Bovy`, there are `r length(get_significant_results(Bovy.model.1))` discoveries. The covariance structure found here is:
```{r, echo=FALSE}
barplot(get_estimated_pi(Bovy.model.1),las = 2, cex.names = 0.7)
```

The row means of the posterior mean matrix are not 0.
```{r}
rowMeans(Bovy.model.1$result$PosteriorMean)[1:6]
```


```{r}
mean((Bovy.model.1$result$PosteriorMean - mashcontrast.model.1.full$result$PosteriorMean)^2)
```

# Compare models

The RRMSE plot:
```{r}
delta.10 = data$C - rowMeans(data$C)
deltahat.10 = data$Chat - rowMeans(data$Chat)

delta.1 = delta.10[, c(2:10, 1)]
deltahat.1 = deltahat.10[, c(2:10, 1)]

barplot(c(sqrt(mean((delta.10 - mashcontrast.model.10.full$result$PosteriorMean)^2)/mean((delta.10 - deltahat.10)^2)), 
          sqrt(mean((delta.10 - Bovy.model.10$result$PosteriorMean)^2)/mean((delta.10 - deltahat.10)^2)),
          sqrt(mean((delta.1 - mashcontrast.model.1.full$result$PosteriorMean)^2)/mean((delta.1 - deltahat.1)^2)),
          sqrt(mean((delta.10 - Bovy.model.1$result$PosteriorMean)^2)/mean((delta.10 - deltahat.10)^2))), ylim=c(0,0.2), names.arg = c('mashcommon.10','Bovy.10', 'mashcommon.1','Bovy.1'), ylab='RRMSE')
```

We check the False Positive Rate and True Positive Rate. 
$$FPR = \frac{|N\cap S|}{|N|} \quad TPR = \frac{|CS\cap S|}{|T|} $$

```{r}
sign.test.mash.10 = as.matrix(delta.10)*mashcontrast.model.10.full$result$PosteriorMean
sign.test.bovy.10 = as.matrix(delta.10)*Bovy.model.10$result$PosteriorMean

sign.test.mash.1 = as.matrix(delta.1)*mashcontrast.model.1.full$result$PosteriorMean
sign.test.bovy.1 = as.matrix(delta.10)*Bovy.model.1$result$PosteriorMean


thresh.seq = seq(0, 1, by=0.0005)[-1]
mashcontrast.10 = matrix(0,length(thresh.seq), 2)
bovy.10 = matrix(0,length(thresh.seq), 2)
colnames(mashcontrast.10) = colnames(bovy.10) = c('TPR', 'FPR')
mashcontrast.1 = matrix(0,length(thresh.seq), 2)
bovy.1 = matrix(0,length(thresh.seq), 2)
colnames(mashcontrast.1) = colnames(bovy.1) = c('TPR', 'FPR')

for(t in 1:length(thresh.seq)){
  mashcontrast.10[t,] = c(sum(sign.test.mash.10>0 & mashcontrast.model.10.full$result$lfsr <= thresh.seq[t])/sum(delta.10!=0), sum(delta.10==0 & mashcontrast.model.10.full$result$lfsr <=thresh.seq[t])/sum(delta.10==0))
  
  bovy.10[t,] = c(sum(sign.test.bovy.10>0& Bovy.model.10$result$lfsr <=thresh.seq[t])/sum(delta.10!=0),  sum(delta.10==0& Bovy.model.10$result$lfsr <=thresh.seq[t])/sum(delta.10==0))
  
  mashcontrast.1[t,] = c(sum(sign.test.mash.1>0 & mashcontrast.model.1.full$result$lfsr <= thresh.seq[t])/sum(delta.1!=0), sum(delta.1==0 & mashcontrast.model.1.full$result$lfsr <=thresh.seq[t])/sum(delta.1==0))
  
  bovy.1[t,] = c(sum(sign.test.bovy.1>0& Bovy.model.1$result$lfsr <=thresh.seq[t])/sum(delta.10!=0),  sum(delta.10==0& Bovy.model.1$result$lfsr <=thresh.seq[t])/sum(delta.10==0))
}
```

```{r, echo=FALSE}
{plot(mashcontrast.10[,'FPR'], mashcontrast.10[,'TPR'], col='red',type='l',xlab = 'FPR', ylab='TPR', main='True Positive vs False Positive')
lines(bovy.10[,'FPR'], bovy.10[,'TPR'], col='cyan')
lines(bovy.1[,'FPR'], bovy.1[,'TPR'], col='green')
lines(mashcontrast.1[,'FPR'], mashcontrast.1[,'TPR'], col='black')
legend('bottomright', c('mashcommon.10','bovy.10','bovy.1','mashcommon.1'),col=c('red', 'cyan', 'green', 'black'),lty=c(1,1,1,1))}
```
