---
title: "Outlier Simulation"
author: "Yuxin Zou"
date: 2018-1-23
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the R version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r R-version, echo=FALSE, results='asis'}
```

```{r, echo=FALSE}
# TEMPORARY.
knitr::opts_chunk$set(eval = TRUE)
```

```{r}
library(mvtnorm); library(mashr)
```

# Generate data
```{r echo=FALSE}
set.seed(10)
nsamp = 500

B.0 = matrix(0,nsamp,2)
b = rnorm(nsamp)
B.1 = cbind(b[1:250],b[1:250])
B.2 = cbind(b[251:500],b[251:500])
b.out = rnorm(nsamp)
B.outlier = cbind(b.out, 2*b.out)

B = rbind(B.0, B.1, B.2, B.outlier)
colnames(B) = c('x','y')

# Add error
Shat = matrix(0.05, nrow = nrow(B), ncol = ncol(B))
E = matrix(rnorm(length(Shat), mean = 0, sd = Shat), nrow = nrow(B),
             ncol = ncol(B))
Bhat = B + E

{plot(Bhat[1:1000,1], Bhat[1:1000,2], xlim=range(Bhat[,1]), ylim=range(Bhat[,2]), xlab = 'x', ylab='y')
points(Bhat[1001:1500,1], Bhat[1001:1500,2], col='red')}
```

Mash with all correct covariance structure:
```{r}
mash_data = mash_set_data(Bhat = Bhat, Shat = Shat)
U.c = cov_canonical(mash_data, cov_methods = 'equal_effects')
U.c = c(U.c, list('inv_effects' = matrix(c(1,2,2,4),2,2)))
mash.model = mash(mash_data, U.c)
barplot(get_estimated_pi(mash.model), las=2, cex.names = 0.7)
length(get_significant_results(mash.model))
```

```{r}
{plot(mash.model$result$PosteriorMean[1:1000,1], mash.model$result$PosteriorMean[1:1000,2])
points(mash.model$result$PosteriorMean[1001:1500,1], mash.model$result$PosteriorMean[1001:1500,2], col='red')}
```

Mash without one covariance structure:

```{r}
mash_data = mash_set_data(Bhat = Bhat, Shat = Shat)
U.c = cov_canonical(mash_data, cov_methods = 'equal_effects')
mash.model.o = mash(mash_data, U.c)
barplot(get_estimated_pi(mash.model.o), las=2, cex.names = 0.7)
length(get_significant_results(mash.model.o))
```

```{r}
which(get_significant_results(mash.model.o) > 1000)
loglik = mash_compute_vloglik(mash.model.o, mash_data)
hist(loglik)
which(loglik < -200)
```

```{r}
{plot(mash.model.o$result$PosteriorMean[1:1000,1], mash.model.o$result$PosteriorMean[1:1000,2])
points(mash.model.o$result$PosteriorMean[1001:1500,1], mash.model.o$result$PosteriorMean[1001:1500,2], col='red')}
```

```{r}
# lower probability
p = numeric(nsamp)
for (i in 1:nsamp){
  p[i] = pmvnorm(lower=-Inf, upper=Bhat[i,], mean=rep(0,2), sigma=rbind(c(1+0.05^2, 1),c(1, 1+0.05^2)))
}
p.small = ifelse(p > 0.5, 1-p,p)

M.dist = numeric(nsamp)
for(i in 1:nsamp){
  M.dist[i] = t(Bhat[i,]) %*% solve(rbind(c(1+0.05^2, 1),c(1, 1+0.05^2))) %*% Bhat[i,]
}

{qqplot(qchisq(ppoints(500), df = 2), M.dist,
       main = expression("Q-Q plot for" ~~ {chi^2}[nu == 2]))
qqline(M.dist, distribution = function(p) qchisq(p, df = 2),
       prob = c(0.1, 0.6), col = 2)}
```

$$\left(\begin{array}{c} \hat{x} \\ \hat{y}\end{array} \right) | \left(\begin{array}{c} x \\ y \end{array} \right) \sim N(\left(\begin{array}{c} x \\ y\end{array} \right), \left( \begin{array}{c c} 0.05^2 & 0 \\
0 & 0.05^2 \end{array} \right))$$
$$\left(\begin{array}{c} x \\ y \end{array} \right) \sim \frac{1}{2} \delta_{0} + \frac{1}{2} N(\left(\begin{array}{c} 0 \\ 0\end{array} \right), \left( \begin{array}{c c} 1 & 1 \\
1 & 1 \end{array} \right))$$

$$\Rightarrow$$ $$\left(\begin{array}{c} \hat{x} \\ \hat{y}\end{array} \right) \sim \frac{1}{2} N(0, \left( \begin{array}{c c} 0.05^2 & 0 \\
0 & 0.05^2 \end{array} \right)) + \frac{1}{2} N(0, \left( \begin{array}{c c} 1+0.05^2 & 1 \\
1 & 1+0.05^2 \end{array} \right))$$ 

```{r}
which(0.5 * dmvnorm(Bhat, mean = rep(0,2), sigma=diag(0.05^2, 2)) + 0.5 * dmvnorm(Bhat, mean = rep(0,2), sigma=rbind(c(1+0.05^2, 1),c(1, 1+0.05^2))))
```

```{r echo=FALSE}
{plot(Bhat[1:1000,1], Bhat[1:1000,2], xlim=range(Bhat[,1]), ylim=range(Bhat[,2]), xlab = 'x', ylab='y')
points(Bhat[1001:1010, 1], Bhat[1001:1010,2], col='red')
points(Bhat[1011:1020,1], Bhat[1011:1020,2], col='blue')}
```

# Identify outlier

We want to identify red points as outliers, but blue points are not.

Firstly, we decompose $\Sigma$ using eigenvalue decomposition, $\Sigma = QDQ'$

```{r}
eigen.sigma = eigen(matrix(c(1,1,1,1),2,2))
```

We can transform the $\Sigma$ to identity matrix, $D^{-1/2}Q' \Sigma = I$.

We apply this transfermation on (x y)'.

```{r}
diag.inv = eigen.sigma$values^{-0.5}
diag.inv[diag.inv == Inf] = 0

B.tran = t(t(B %*% eigen.sigma$vectors) * diag.inv)
```

```{r echo=FALSE}
{plot(B.tran[,1], B.tran[,2], xlim=range(B.tran[,1]), ylim=range(B.tran[,2]), xlab = 'x', ylab='y')
points(Bhat.tran[1001:1010, 1], Bhat.tran[1001:1010,2], col='red')
points(Bhat.tran[1011:1020,1], Bhat.tran[1011:1020,2], col='blue')}
```

# Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
