---
title: "Outlier Simulation"
author: "Yuxin Zou"
date: 2018-6-20
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r echo=FALSE}
library(mvtnorm); library(mashr)
```

# Generate data

$$\left(\begin{array}{c} \hat{\beta}_{j1} \\ \hat{\beta}_{j2} \end{array} \right) | \left(\begin{array}{c} \beta_{j1} \\ \beta_{j2}\end{array} \right) \sim N(\left(\begin{array}{c} \beta_{j1} \\ \beta_{j2}\end{array} \right), \left( \begin{array}{c c} 0.05^2 & 0 \\
0 & 0.05^2 \end{array} \right)) $$
$$\left(\begin{array}{c} \beta_{j1} \\ \beta_{j2}\end{array} \right) \sim \frac{1}{2} \delta_{0} + \frac{1}{2}N(0, \left( \begin{array}{c c} 1 & 1 \\
1 & 1 \end{array} \right))$$

```{r echo=FALSE}
set.seed(10)
nsamp = 500

B.0 = matrix(0,nsamp,2)
b = rnorm(nsamp)
B.1 = cbind(b, b)

B = rbind(B.0, B.1)
colnames(B) = c('x','y')

# Add error
Shat = matrix(0.05, nrow = nrow(B), ncol = ncol(B))
E = matrix(rnorm(length(Shat), mean = 0, sd = Shat), nrow = nrow(B),
             ncol = ncol(B))
Bhat = B + E

plot(Bhat[1:1000,1], Bhat[1:1000,2], xlim=range(Bhat[,1]), ylim=range(Bhat[,2]), xlab = 'x', ylab='y')

```

Fit `mash` model:
```{r echo=FALSE}
mash_data = mash_set_data(Bhat = Bhat, Shat = Shat)
U.c = cov_canonical(mash_data, cov_methods = 'equal_effects')
mash.model = mash(mash_data, U.c)
barplot(get_estimated_pi(mash.model), las=2, cex.names = 0.7)
```
The identified covariance structures are correct.

```{r}
est.lower = mash_data$Bhat+1.96*mash_data$Shat > mash.model$result$PosteriorMean-1.96*mash.model$result$PosteriorSD
est.upper = mash_data$Bhat-1.96*mash_data$Shat < mash.model$result$PosteriorMean+1.96*mash.model$result$PosteriorSD
est = est.lower * est.upper
estSum = rowSums(est)

mash_plot_meta(mash.model, 838)

library(ggplot2)
gene = data.frame(mash_data$Bhat[838,])
colnames(gene) = 'EffectSize'
gene$Group = row.names(gene)
gene$se = data.frame(mash_data$Shat[838,])
ggplot(gene, aes(y = EffectSize, x = Group, color=Group)) + 
  geom_point() +
  geom_errorbar(aes(ymin=EffectSize-1.96*se, ymax=EffectSize+1.96*se), width=0.4) + 
  theme_bw(base_size=12) + coord_flip()
```

We simulate a new dataset.
$$\left(\begin{array}{c} \hat{\beta}_{j1} \\ \hat{\beta}_{j2} \end{array} \right) | \left(\begin{array}{c} \beta_{j1} \\ \beta_{j2}\end{array} \right) \sim N(\left(\begin{array}{c} \beta_{j1} \\ \beta_{j2}\end{array} \right), \left( \begin{array}{c c} 0.05^2 & 0 \\
0 & 0.05^2 \end{array} \right)) $$
$$\left(\begin{array}{c} \beta_{j1} \\ \beta_{j2}\end{array} \right) \sim \frac{1}{3} \delta_{0} + \frac{1}{3}N(0, \left( \begin{array}{c c} 1 & 1 \\
1 & 1 \end{array} \right)) + \frac{1}{3}N(0, \left( \begin{array}{c c} 1 & -1 \\
-1 & 1 \end{array} \right))$$

```{r}
set.seed(20)
nsamp = 500

B.0 = matrix(0,nsamp,2)
b = rnorm(nsamp)
B.1 = cbind(b, b)
b.out = rnorm(nsamp)
B.outlier = cbind(b.out, -b.out)

B = rbind(B.0, B.1, B.outlier)
colnames(B) = c('x','y')

# Add error
Shat = matrix(0.05, nrow = nrow(B), ncol = ncol(B))
E = matrix(rnorm(length(Shat), mean = 0, sd = Shat), nrow = nrow(B),
             ncol = ncol(B))
Bhat = B + E

{plot(Bhat[1:1000,1], Bhat[1:1000,2], xlim=range(Bhat[,1]), ylim=range(Bhat[,2]), xlab = 'x', ylab='y')
points(Bhat[1001:1500,1], Bhat[1001:1500,2], col='red')
}
```

If we fit `mash` model on this new data, without the correct covariance structure:

```{r}
mash.data.o = mash_set_data(Bhat = Bhat, Shat=Shat)
U.c = cov_canonical(mash.data.o, cov_methods = 'equal_effects')
mash.model.o = mash(mash.data.o, U.c)
barplot(get_estimated_pi(mash.model.o), las=2, cex.names = 0.7)
# mash.model.o = mash.model
# posterior = mash_compute_posterior_matrices(mash.model, mash.data.o)
# mash.model.o$result = posterior
length(get_significant_results(mash.model.o))
```

```{r}
{plot(mash.model.o$result$PosteriorMean[1:1000,1], mash.model.o$result$PosteriorMean[1:1000,2])
points(mash.model.o$result$PosteriorMean[1001:1500,1], mash.model.o$result$PosteriorMean[1001:1500,2], col='red')}
```
The outlier points are all shrinkaged close to 0. We want to know whether the covariance structures are sufficient before using `mash` model.

# Identify points

```{r}
est.lower = mash.data.o$Bhat+1.96*mash.data.o$Shat > mash.model.o$result$PosteriorMean-1.96*mash.model.o$result$PosteriorSD
est.upper = mash.data.o$Bhat-1.96*mash.data.o$Shat < mash.model.o$result$PosteriorMean+1.96*mash.model.o$result$PosteriorSD
est = est.lower * est.upper
estSum = rowSums(est)
summary(estSum)
```
```{r}
cov(mash.data.o$Bhat[estSum == 0,])
```

# Is there a way that identify the problem before fitting the mash model?

Make points have max norm 1
```{r}
Bhatnorm = apply(mash.data.o$Bhat, 1, function(x) sqrt(sum(x^2)))
Bhat = mash.data.o$Bhat/max(Bhatnorm)
Shat = mash.data.o$Shat/(max(Bhatnorm)^2)
```

```{r}
{plot(Bhat[1:1000,1],Bhat[1:1000,2])
points(Bhat[1001:1500,1],Bhat[1001:1500,2])}
```

```{r}
mash.data = mash_set_data(Bhat, Shat)
Cov = mashr:::expand_cov(U.c, 1, TRUE)
lm <- mashr:::calc_relative_lik_matrix(mash.data, Cov, algorithm.version='Rcpp')
rowMax = apply(lm$loglik_matrix, 1, max)
```

# Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
