---
title: "Simulation with Signal - Compare miximash"
author: "Yuxin Zou"
date: 2018-05-19
output: 
  html_document:
    code_folding: hide
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

```{r code-version, echo=FALSE, results='asis'}
```

```{r}
library(mashr); library(miximash)
source('../code/plotCormotif.R')
```

# Simulation Design

In the simulation, $$c_{j2...R} = c_{j1}1 + \delta_{j}$$
$$\hat{c}_{j}|c_{j} \sim N_{R}(c_{j}, \frac{1}{2}I)$$
Let L be the contrast matrix. Therefore, $$\hat{\delta}_{j}|\delta_{j} = L\hat{c}_{j}|c_{j} \sim N_{R-1}(\delta_{j}, \frac{1}{2}LL')$$

We first generate the data:
```{r}
set.seed(2018)
data = sim_contrast2(nsamp = 12000, ncond = 8)
```
This simulation routine creates a dataset with 8 conditions, and four different types of deviations $\delta_{j}$: null, independent among conditions, condition-specific in condition 2, and shared (equal effects in all conditions). The data contains 10% non-null effects 
$$\delta_{j} \sim \frac{9}{10}N(0,0)+\frac{1}{30}N(0,I)+\frac{1}{30}N(0,11^{T})+\frac{1}{30}N(0,e_{1}e_{1}^{T})$$

Set up the contrast matrix and the mash contrast data object
```{r}
R = 8
L = diag(R-1)
L = cbind(rep(-1, R-1), L)
row.names(L) = c('2-1','3-1','4-1','5-1','6-1','7-1','8-1')

mash_data = mash_set_data(Bhat=data$Chat, Shat=data$Shat)
mash_data_L = mash_set_data_contrast(mash_data, L)
```

# Mashcommonbaseline
Set up the covariance matrices:

```{r}
# canonical
U.c = cov_canonical(mash_data_L)
# data driven
m.1by1 = mash_1by1(mash_data_L, alpha=0)
strong = get_significant_results(m.1by1,0.05)
# only 1 strong samples
```

Fit `mashcontrast` model
```{r}
mashcontrast.model = mash(mash_data_L, U.c, algorithm.version = 'R')
```

The log likelihood is
```{r}
print(get_loglik(mashcontrast.model),digits = 10)
```

Use `get_significant_results` to find the indices of effects that are “significant”:

```{r}
length(get_significant_results(mashcontrast.model))
```

The number of false positive is `r sum(get_significant_results(mashcontrast.model) < 12000-1200)`. 

```{r}
barplot(get_estimated_pi(mashcontrast.model),las = 2,cex.names = 0.7)
```

# Mash
```{r}
Indep.data = mash_set_data(mash_data_L$Bhat, mash_data_L$Shat)
Indep.m = mash(Indep.data, U.c)
```

The log likelihood is
```{r}
print(get_loglik(Indep.m),digits = 10)
```

Use `get_significant_results` to find the indices of effects that are “significant”:

```{r}
length(get_significant_results(Indep.m))
```

The number of false positive is `r sum(get_significant_results(Indep.m) < 12000-1200)`. 

```{r}
barplot(get_estimated_pi(Indep.m),las = 2,cex.names = 0.7)
```

# Miximash

```{r}
fit = miximash(mash_data_L$Bhat, mash_data_L$Shat, mess = TRUE, K = 2:6)
```

Loglikelihood
```{r}
plot(2:6, fit$loglike[,2],type = "l",xlab = "K",ylab = "loglike")
```

```{r}
plotMotif(fit$allmotif[[2]])
```

Discoveries: 
```{r}
sum(rowSums(fit$lfsr[[1]] <= 0.05) > 0)
```

# Compare

The RRMSE plot:
```{r}
delta = data$C %*% t(L)
barplot(c(sqrt(mean((delta - m.1by1$result$PosteriorMean)^2)/mean((delta - data$Chat%*%t(L))^2)), 
          sqrt(mean((delta - mashcontrast.model$result$PosteriorMean)^2)/mean((delta - data$Chat%*%t(L))^2)), 
          sqrt(mean((delta - Indep.m$result$PosteriorMean)^2)/mean((delta - data$Chat%*%t(L))^2)), 
          sqrt(mean((delta - fit$post_mean[[1]])^2)/mean((delta - data$Chat%*%t(L))^2))), ylim=c(0,0.8), names.arg = c('ash','mashcommonbaseline', 'mashIndep', 'miximash'), ylab='RRMSE')
```

We check the False Positive Rate and True Positive Rate. 
$$FPR = \frac{|N\cap S|}{|N|} \quad TPR = \frac{|CS\cap S|}{|T|} $$

Each effect is treated as a separate discovery in each condition
```{r}
delta = data$C %*% t(L)

sign.test.mash = as.matrix(delta)*mashcontrast.model$result$PosteriorMean
sign.test.Indep = as.matrix(delta)*Indep.m$result$PosteriorMean
sign.test.ash = as.matrix(delta)*m.1by1$result$PosteriorMean
sign.test.mix = as.matrix(delta)*fit$post_mean[[1]]

thresh.seq = seq(0, 1, by=0.0005)[-1]
mashcontrast = matrix(0,length(thresh.seq), 2)
Indep = matrix(0,length(thresh.seq), 2)
Ash = matrix(0,length(thresh.seq), 2)
Mix = matrix(0,length(thresh.seq), 2)
colnames(mashcontrast) = colnames(Indep) = colnames(Ash) = colnames(Mix) = c('TPR', 'FPR')

for(t in 1:length(thresh.seq)){
  mashcontrast[t,] = c( sum(sign.test.mash>0 & mashcontrast.model$result$lfsr <= thresh.seq[t])/sum(delta!=0), sum(delta==0 & mashcontrast.model$result$lfsr <=thresh.seq[t])/sum(delta==0))
  
  Indep[t,] = c(sum(sign.test.Indep>0& Indep.m$result$lfsr <=thresh.seq[t])/sum(delta!=0),  sum(delta==0& Indep.m$result$lfsr <=thresh.seq[t])/sum(delta==0))
  
  Ash[t,] = c(sum(sign.test.ash>0& m.1by1$result$lfsr <=thresh.seq[t])/sum(delta!=0),  sum(delta==0& m.1by1$result$lfsr <=thresh.seq[t])/sum(delta==0))
  
  Mix[t,] = c(sum(sign.test.mix>0& fit$lfsr[[1]] <=thresh.seq[t])/sum(delta!=0),  sum(delta==0& fit$lfsr[[1]] <=thresh.seq[t])/sum(delta==0))
}

```

```{r}
{plot(mashcontrast[,'FPR'], mashcontrast[,'TPR'], col='red',type='l',ylab="True Positive Rate",xlab="False Positive Rate")
lines(Indep[,'FPR'], Indep[,'TPR'])
lines(Ash[,'FPR'], Ash[,'TPR'], col='green')
lines(Mix[,'FPR'], Mix[,'TPR'], col='cyan')
legend('bottomright', c('mashcommonbaseline', 'mashIndep', 'ash', 'miximash'),col=c('red','black','green','cyan'),lty=c(1,1,1,1))
}
```

# Session information

```{r session-info}
```
