---
title: "Simulation with Signal Example"
author: "Yuxin Zou"
date: 2018-01-11
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the R version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r R-version, echo=FALSE, results='asis'}
```

```{r, echo=FALSE}
# TEMPORARY.
knitr::opts_chunk$set(eval = TRUE)
```

```{r, echo=FALSE}
library(ashr)
library(plyr)
library(mvtnorm)
library(assertthat)
source('~/Documents/GitHub/mashr-zou/R/compute_covs.R')
source('~/Documents/GitHub/mashr-zou/R/cov_udi.R')
source('~/Documents/GitHub/mashr-zou/R/data2cov.R')
source('~/Documents/GitHub/mashr-zou/R/ed.R')
source('~/Documents/GitHub/mashr-zou/R/est_cor.R')
source('~/Documents/GitHub/mashr-zou/R/get_functions.R')
source('~/Documents/GitHub/mashr-zou/R/likelihoods_origdata.R')
source('~/Documents/GitHub/mashr-zou/R/likelihoods_scaleddata.R')
source('~/Documents/GitHub/mashr-zou/R/mash.R')
source('~/Documents/GitHub/mashr-zou/R/opt.R')
source('~/Documents/GitHub/mashr-zou/R/plots.R')
source('~/Documents/GitHub/mashr-zou/R/posterior.R')
source('~/Documents/GitHub/mashr-zou/R/posterior_common_cov.R')
source('~/Documents/GitHub/mashr-zou/R/posterior_lowmem.R')
source('~/Documents/GitHub/mashr-zou/R/RcppExports.R')
source('~/Documents/GitHub/mashr-zou/R/set_data.R')
source('~/Documents/GitHub/mashr-zou/R/simulations.R')
```

# Simulation Design

In the simulation, $$c_{j2...R} = c_{j1} + \delta_{j}$$
$$\hat{c}_{j}|c_{j} \sim N_{R}(c_{j}, \frac{1}{2}I)$$
Let L be the contrast matrix. Therefore, $$\hat{\delta}_{j}|\delta_{j} = L\hat{c}_{j}|c_{j} \sim N_{R-1}(\delta_{j}, \frac{1}{2}LL')$$

We first generate the data:
```{r}
set.seed(2018)
data = sim2(nsamp = 21000, ncond = 8)
```
This simulation routine creates a dataset with 8 conditions, and four different types of deviations $\delta_{j}$: null, independent among conditions, condition-specific in condition 2, and shared (equal effects in all conditions). The data contains 10% non-null effects 
$$\delta_{j} \sim \frac{9}{10}N(0,0)+\frac{1}{30}N(0,I)+\frac{1}{30}N(0,11^{T})+\frac{1}{30}N(0,e_{1}e_{1}^{T})$$

Set up the contrast matrix and the mash contrast data object
```{r}
R = 8
L = diag(R-1)
L = cbind(rep(-1, R-1), L)
row.names(L) = c('2-1','3-1','4-1','5-1','6-1','7-1','8-1')

mash_data = mash_set_data(Bhat=data$Chat, Shat=data$Shat)
mash_data_L = mash_set_data_contrast(mash_data, L)
```

Set up the covariance matrices:

```{r}
# canonical
U.c = cov_canonical(mash_data_L)
# data driven
m.1by1 = mash_1by1(mash_data_L, alpha=1)
strong = get_significant_results(m.1by1,0.05)
# only 5 strong samples

# center Z
# mash_data_L.center = mash_data_L
# mash_data_L.center$Bhat = mash_data_L$Bhat/mash_data_L$BShat # obtain z
# mash_data_L.center$BShat = matrix(1, nrow(mash_data_L$Bhat),ncol(mash_data_L$Bhat))
# mash_data_L.center$Bhat = apply(mash_data_L.center$Bhat, 2, function(x) x - mean(x))
# U.pca = cov_pca(mash_data_L.center,5,strong)
# U.ed = cov_ed(mash_data_L.center, U.pca, strong)
```

Fit `mashcontrast` model
```{r}
mashcontrast.model = mash(mash_data_L, U.c, algorithm.version = 'R')
print(get_loglik(mashcontrast.model),digits = 10)
```

Use `get_significant_results` to find the indices of effects that are “significant”:

```{r}
length(get_significant_results(mashcontrast.model))
```

The number of false positive is `r sum(get_significant_results(mashcontrast.model) < 18000-1800)`.

```{r}
barplot(get_estimated_pi(mashcontrast.model),las = 2,cex.names = 0.7)
```
<!--
The weights are mainly in the canonical matrices, as we expected. The data is generated using the canonical matrices.
-->

# Comparing discovery result with $\hat{\delta}_{j} \sim N_{7}(\delta_{j},I)$:

```{r}
Indep.data = mash_set_data(Bhat = data$Chat%*%t(L))
Indep.model = mash(Indep.data, U.c, algorithm.version = 'R')
print(get_loglik(Indep.model),digits = 10)
```
```{r}
length(get_significant_results(Indep.model))
```

There are more discoveries here, but lots of them are false discoveries. The number of false positive is `r sum(get_significant_results(Indep.model) < 18000-1800)`.

```{r}
barplot(get_estimated_pi(Indep.model),las = 2, cex.names = 0.7)
```

<!--
We check the False Positive Rate and True Positive Rate. 
$$FPR = \frac{|N\cap S|}{|N|} \quad TPR = \frac{|CS\cap S|}{|T|} $$

```{r, eval=FALSE, echo=FALSE}
CS_S = function(model, thresh=0.05, data){
  sig.index = model$result$lfsr <= thresh
  sum(sig.index * model$result$PosteriorMean * data > 0)
}

N_S = function(model, thresh=0.05, data){
  N.index = data == 0
  sig.index = model$result$lfsr <= thresh
  sum(sig.index * N.index)
}

delta = data$C %*% t(L)
N = sum(delta == 0)
Tr = nrow(delta) * ncol(delta) - N

thresh.seq = seq(0, 0.5, by=0.0005)[-1]
mashcontrast = matrix(0,length(thresh.seq), 2)
Indep = matrix(0,length(thresh.seq), 2)
colnames(mashcontrast) = c('TPR', 'FPR')
colnames(Indep) = c('TPR', 'FPR')
for(t in 1:length(thresh.seq)){
  mashcontrast[t,] = c(CS_S(mashcontrast.model, thresh.seq[t], delta)/Tr, N_S(mashcontrast.model, thresh.seq[t],delta)/N)
  
  Indep[t,] = c(CS_S(Indep.model, thresh.seq[t], delta)/Tr,  N_S(Indep.model, thresh.seq[t],delta)/N)
}
```

```{r}
# plot(Indep[,'FPR'], Indep[,'TPR'], xlab = 'FPR', ylab='TPR', type='l', xlim=c(0,0.8), ylim=c(0,0.6))
# lines(mashcontrast[,'FPR'], mashcontrast[,'TPR'], col='red')
```

Or
$$FPR = \frac{|S\cap N|}{|N|} \quad TPR = \frac{|S|}{|T|} $$
```{r}
delta = data$C %*% t(L)
Tr = sum(apply(delta!=0, 1, max)) # number of lines with at least one non-zero effects
N = nrow(delta) - Tr # number of lines with all zero effects

thresh.seq = seq(0, 0.5, by=0.0005)[-1]
mashcontrast = matrix(0,length(thresh.seq), 2)
Indep = matrix(0,length(thresh.seq), 2)
colnames(mashcontrast) = c('TPR', 'FPR')
colnames(Indep) = c('TPR', 'FPR')

for(t in 1:length(thresh.seq)){
  mashcontrast[t,] = c(length(get_significant_results(mashcontrast.model, thresh = thresh.seq[t]))/Tr, sum(get_significant_results(mashcontrast.model, thresh = thresh.seq[t])<=16200)/N)
  Indep[t,] = c(length(get_significant_results(Indep.model, thresh = thresh.seq[t]))/Tr, sum(get_significant_results(Indep.model, thresh = thresh.seq[t])<=16200)/N)
}

# plot(mashcontrast[,'FPR'], mashcontrast[,'TPR'], xlab = 'FPR', ylab='TPR', type='l')
# lines(Indep[,'FPR'], Indep[,'TPR'], col='red')
```
-->

# Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
