---
title: "Simulation with Signal Example"
author: "Yuxin Zou"
date: 2018-01-10
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the R version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r R-version, echo=FALSE, results='asis'}
```

```{r, echo=FALSE}
# TEMPORARY.
knitr::opts_chunk$set(eval = TRUE)
```

Loading the functions:

```{r}
library(ashr)
library(plyr)
library(mvtnorm)
library(assertthat)
source('~/Documents/GitHub/mashr-zou/R/compute_covs.R')
source('~/Documents/GitHub/mashr-zou/R/cov_udi.R')
source('~/Documents/GitHub/mashr-zou/R/data2cov.R')
source('~/Documents/GitHub/mashr-zou/R/ed.R')
source('~/Documents/GitHub/mashr-zou/R/est_cor.R')
source('~/Documents/GitHub/mashr-zou/R/get_functions.R')
source('~/Documents/GitHub/mashr-zou/R/likelihoods_origdata.R')
source('~/Documents/GitHub/mashr-zou/R/likelihoods_scaleddata.R')
source('~/Documents/GitHub/mashr-zou/R/mash.R')
source('~/Documents/GitHub/mashr-zou/R/opt.R')
source('~/Documents/GitHub/mashr-zou/R/plots.R')
source('~/Documents/GitHub/mashr-zou/R/posterior.R')
source('~/Documents/GitHub/mashr-zou/R/posterior_common_cov.R')
source('~/Documents/GitHub/mashr-zou/R/posterior_lowmem.R')
source('~/Documents/GitHub/mashr-zou/R/RcppExports.R')
source('~/Documents/GitHub/mashr-zou/R/set_data.R')
source('~/Documents/GitHub/mashr-zou/R/simulations.R')
```

# Simulation Design

In the simulation, $$c_{j2...8} = c_{j1} + \delta_{j}$$
$$\hat{c}_{j} \sim N_{8}(c_{j}, \frac{1}{2}I)$$
Let L be the contrast matrix. Therefore, $$\hat{\delta}_{j} = L\hat{c}_{j} \sim N_{7}(\delta_{j}, \frac{1}{2}LL')$$

We first generate the data:
```{r}
set.seed(1)
data = sim2(nsamp = 2000, ncond = 8)
```
This simulation routine creates a dataset with 8 conditions, and four different types of deviations $\delta_{j}$: null, independent among conditions, condition-specific in condition 2, and shared (equal effects in all conditions). It creates 2000 effects of each type for a total of 8000 effects.

Set up the contrast matrix and the mash contrast data object
```{r}
L = rbind(c(-1,1,0,0,0,0,0,0),
          c(-1,0,1,0,0,0,0,0),
          c(-1,0,0,1,0,0,0,0),
          c(-1,0,0,0,1,0,0,0),
          c(-1,0,0,0,0,1,0,0),
          c(-1,0,0,0,0,0,1,0),
          c(-1,0,0,0,0,0,0,1))
row.names(L) = c('2-1','3-1','4-1','5-1','6-1','7-1','8-1')
mash_data = mash_set_data(Bhat=data$Chat, Shat=data$Shat)
mash_data_L = mash_set_data_contrast(mash_data, L)
```

Set up the covariance matrices:

```{r}
U.c = cov_canonical(mash_data_L)
```

Fit `mashcontrast` model
```{r}
mashcontrast.model = mash(mash_data_L, U.c, algorithm.version = 'R')
```

Use `get_significant_results` to find the indices of effects that are “significant”:

```{r}
length(get_significant_results(mashcontrast.model))
```

```{r}
barplot(get_estimated_pi(mashcontrast.model),las = 2)
```

```{r}
print(get_loglik(mashcontrast.model),digits = 10)
```

# Comparing discover result with $\hat{\delta}_{j} \sim N_{7}(\delta_{j},I)$:

```{r}
Indep.data = mash_set_data(Bhat = data$Chat%*%t(L))
U.c = cov_canonical(Indep.data)
Indep.model = mash(Indep.data, U.c, algorithm.version = 'R')
```
```{r}
length(get_significant_results(Indep.model))
```
```{r}
barplot(get_estimated_pi(Indep.model),las = 2)
```


# Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
